---
title: "Reproducible Research: Peer Assessment 1"
author: "Christel Veefkind"
date: "27 juli 2016"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data


```{r cars}
  library(ggplot2)
  setwd("~/Coursera/RepData")
  activity <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
```{r activity, echo=TRUE}
      sum(activity$steps,na.rm =TRUE)
```
## Histogram of total number of steps taken per day

```{r stepsPerDay, echo=TRUE}
      stepsPerDay <- aggregate(activity$steps, list(date = activity$date),sum)
      hist(stepsPerDay$x)
```

## Mean and Median of the total number of steps taken per day
```{r Mean, echo=TRUE}
      summary(stepsPerDay$x)
```
## Average daily activity pattern

```{r daily pattern, echo=TRUE}
      stepsPerInterval <- aggregate(activity$steps, list(interval = activity$interval),mean, na.rm = TRUE)
      plot(stepsPerInterval$interval, stepsPerInterval$x, type = "l")
      IntervalHighestSteps <- subset(stepsPerInterval, x == max(stepsPerInterval$x), select = interval)
      View(IntervalHighestSteps)
```
## Missing values
 

```{r na_steps, echo=TRUE}
      na_steps <- activity[rowSums(is.na(activity)) > 0,]
      nrow(na_steps)
      
    ### Filling in missing values with average per interval
      activity_filled <- activity
      activity_filled$steps[is.na(activity_filled$steps)] <- round(stepsPerInterval$x)
    ###  Steps per day with NA values filled with the mean per interval
      stepsPerDay_filled <- aggregate(activity_filled$steps, list(date = activity_filled$date),mean)
      hist(stepsPerDay_filled$x)
      
      summary(stepsPerDay_filled$x)
``` 
There is no difference in mean and median

## Are there differences in activity patterns between weekdays and weekends?

```{r weekday, echo=TRUE}
      activity_filled$day <- weekdays(as.Date(activity_filled$date))
      activity_filled_grouped <- aggregate(activity_filled$steps, by=list(activity_filled$interval, activity_filled$day), FUN=mean)
      activity_filled_grouped$dayType <- "weekday"
      activity_filled_grouped$dayType[activity_filled_grouped$Group.2 %in% c("zaterdag","zondag")] <- "weekend"
      activity_filled_grouped$dayType <- as.factor(activity_filled_grouped$dayType)
        qplot(Group.1,x , data = activity_filled_grouped, 
        geom = "line",
        xlab =  "interval",
        ylab =  "average number of steps",
        main = "Average steps Weekend versus Weekdays",
        facets =dayType ~.)
      
